Left-linear rules in adhesive categories

### 2024-02-11 icalp/fusione.tex
minor things

% direct shift
\newcommand{\shiftdir}[1][]{\ensuremath{\mathrel{{\rightsquigarrow}^{\mathit{sh}}_{#1}}}}

@@

\begin{definition}[Switch equivalence]
  \label{de:switch-equivalence}
  Let $(\X, \R)$ be a left-linear DPO rewriting system.  Let
  $\der{D}, \der{E} : G \Mapsto H$ be derivations with the same
  length, $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ and
  $\der{E}=\{\dder{E}_{i}\}_{i=0}^n$. If
  $\dder{D}_i \cdot \dder{D}_{i+1}$ is a switch of
  $\dder{E}_i \cdot \dder{E}_{i+1}$ for some $i \in [0,n-1]$ we write
  $\der{D} \leftrightsquigarrow_{(i,i+1)} \der{E}$. Then \emph{switch
    equivalence} is the least equivalence on derivation such that if
  $\der{D} \leftrightsquigarrow_{(i,i+1)} \der{E}$ then
  $\der{D} \equiv^s \der{E}$.
\end{definition}

@@

The characterisation of switch equivalence based on consistent
permutation for linear rules allows to derive some simple but
fundamental results for the theory, namely
\begin{itemize}
\item the result of switching does not depends on the order, but only
  on the permutation;
\item independence is global, meaning that if two independent steps are
  moved forward or backward in a derivation, they remain
  independent.
\end{itemize}

@@

As a beginning step in our analysis of switch equivalence, we will
establish three lemmas dealing with derivations of length
$3$. \todo{We will probably limit tho this}

@@

\todo{dividere in due il lemma dopo. devono venire tre lemmi}
\begin{restatable}[Three steps Lemma]{lemma}{lemIndepGlobalLeft}
  \label{lem:indep-global-left}
  Let $\X$ be an $\mathcal{M}$-adhesive category and
  $(\X,\R)$ a left-linear DPO rewriting system.
  %
  Consider a derivation
  $\der{D}=\{\dder{D}_i\}_{i=0}^2$ and suppose that $(i_0,i_1)$ is a
  good independence pair between $\dder{D}_0$ and $\dder{D}_1$,
  $(a_0,a_1)$ one between $\dder{D}_1$ and $\dder{D}_2$ and
  $(e_0, e_1)$ one between $\dder{D}_0$ and
  $S_{a_0,a_1}(\dder{D}_2)$. Then the following properties hold true.
  \begin{enumerate}
  \item $S_{e_0,e_1}(\dder{D}_0)$ and $S_{a_0,a_1}(\dder{D}_1)$ are
     sequentially independent.
  \item If $S_{i_0, i_1}(\dder{D}_0)\updownarrow_! \dder{D}_2$ with a
    good independence pair $(\alpha_0, \alpha_1)$, then
    $S_{i_0,i_1}(\dder{D}_1)$ and $S_{\alpha_0, \alpha_1}(\dder{D}_2)$
    are  sequentially independent.
  \end{enumerate}
\end{restatable}

>>>>

% direct shift
\newcommand{\shift}[1][]{\ensuremath{\mathrel{{\leftrightsquigarrow}^{\mathit{sh}}_{#1}}}}

@@

\begin{definition}[Switch equivalence]
  \label{de:switch-equivalence}
  Let $(\X, \R)$ be a left-linear DPO rewriting system.  Let
  $\der{D}, \der{E} : G \Mapsto H$ be derivations with the same
  length, $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ and
  $\der{E}=\{\dder{E}_{i}\}_{i=0}^n$. If
  $\dder{D}_i \cdot \dder{D}_{i+1}$ is a switch of
  $\dder{E}_i \cdot \dder{E}_{i+1}$ for some $i \in [0,n-1]$ we write
  $\der{D} \shift_{(i,i+1)} \der{E}$. Then \emph{switch
    equivalence} is the least equivalence on derivation such that if
  $\der{D} \leftrightsquigarrow_{(i,i+1)} \der{E}$ then
  $\der{D} \equiv^s \der{E}$.
\end{definition}

@@

In the case of linear rewriting systems, the characterisation of switch equivalence based on consistent
permutation allows to derive some simple but
fundamental results for the theory, namely
\begin{itemize}
\item the result of switching does not depends on the order, but only
  on the permutation;
\item independence is global, meaning that if two independent steps are
  moved forward or backward in a derivation, they remain
  independent.
\end{itemize}

@@

As a beginning step in our analysis of switch equivalence, we will
establish three lemmas dealing with derivations of length
$3$.

@@

\todo{dividere in due il lemma dopo. devono venire tre lemmi}
\begin{restatable}[Second 3-steps Lemma]{lemma}{lemIndepGlobalLeft}
  \label{lem:indep-global-left}
  Let $\X$ be an $\mathcal{M}$-adhesive category and
  $(\X,\R)$ a left-linear DPO rewriting system.
  %
  Consider a derivation
  $\der{D}=\{\dder{D}_i\}_{i=0}^2$ and suppose that $(i_0,i_1)$ is a
  good independence pair between $\dder{D}_0$ and $\dder{D}_1$,
  $(a_0,a_1)$ one between $\dder{D}_1$ and $\dder{D}_2$ and
  $(e_0, e_1)$ one between $\dder{D}_0$ and
  $S_{a_0,a_1}(\dder{D}_2)$. Then the following properties hold true.
  \begin{enumerate}
  \item $S_{e_0,e_1}(\dder{D}_0)$ and $S_{a_0,a_1}(\dder{D}_1)$ are
     sequentially independent.
  \item If $S_{i_0, i_1}(\dder{D}_0)\updownarrow_! \dder{D}_2$ with a
    good independence pair $(\alpha_0, \alpha_1)$, then
    $S_{i_0,i_1}(\dder{D}_1)$ and $S_{\alpha_0, \alpha_1}(\dder{D}_2)$
    are  sequentially independent.
  \end{enumerate}
\end{restatable}
